Tropical mathematics and the Lambda-calculus I

### 2023-07-30 CSL24/section2bisbis.tex
Commit
Dettagli vari

Let us start with a very basic probabilistic language:
the terms are $M::= \true \mid \false \mid M\oplus_p M$, for $p\in[0,1]$, and the operational semantics is $M\oplus_p N\to pM$ and $M\oplus_p N \to (1-p)N$, so that $M\oplus_p N$ plays the role of a probabilistic coin toss of bias $p$.
Consider the program
$$
 M:=(\true \oplus_p\false)\oplus_p((\true\oplus_p\false)\oplus_p(\false\oplus_p\true)).
 $$
 Calling $q=1-p$, to each occurrence of $\true$ or $\false$ in $M$, univocally determined by an address
$\omega\in \{l,r\}^{*}$, is associated a monomial $P_{\omega}(p,q)$ which determines the probability of the event ``$M\twoheadrightarrow_{\omega} \true/\false$'', that is, that $M$ reduces to $\true/\false$ according to the choices in $\omega$.
Thinking of $p,q$ as parameters, $P_{\omega}(p,q)$ can thus be read as the \emph{likelihood function} of the event ``$M\twoheadrightarrow \true_\omega/\false_\omega$''.
 For instance, we have
$P_{rll}(p,q):=qp^2$,
$P_{rrr}(p,q):=q^3$, and 
$P_{rrl}(p,q)=P_{rlr}(p,q):=q^2p$.
The polynomial function $Q_{\true}(p,q):=P_{ll}(p,q)+P_{rll}(p,q)+P_{rrr}(p,q)=p^2+p^2q+q^3$ gives instead the probability of the event ``$M\twoheadrightarrow \true$'', and analogously for $Q_{\false}(p,q):=P_{lr}(p,q)+P_{rrl}(p,q)+P_{rlr}(p,q)=pq+2pq^2$.
This way, the probabilistic evaluation of $M$ is presented as a \emph{hidden Markov model} \cite{Baumr966}, a fundamental statistical model, and notably one to which tropical methods are generally applied \cite{Pachter2ll4}.
Typical questions in this case would be, for a fixed $\omega_0$:
%
%The tropical point of view allows now to express two natural questions about this situation:
\begin{enumerate}
 \item Which is the \emph{maximum likelihood estimator} for the event ``$M\twoheadrightarrow \false_{\omega_0}$''?
 I.e., which is the choice of $p,q$ that maximizes the probability $P_{\omega_0}$?
 \item 
Knowing that $M$ produced $\true$ (similarly for $\false$) , which is the \emph{maximum likelihood estimator} for the event ``$M\twoheadrightarrow \false_{\omega_0}$'', knowing that ``$M\twoheadrightarrow \false$''?
I.e., which is the choice of $p,q$ that makes $\omega_0$ the most likely path among those leading to $\false$ (i.e.\ that maximizes the conditional probability $\BB P(``M\twoheadrightarrow \false_{\omega_0}'' \mid ``M\twoheadrightarrow \false'')$)?
\end{enumerate}

@@

This example is inspired from \cite{Manzo2013}. Again, we start from a very basic non-deterministic language with terms $M::= \true \mid  \mathsf{gen} \mid  M + M$, with an operation semantics comprising a non-deterministic reduction rule 
$M_{1}+M_{2} \stackrel{\alpha}{\to} M_{i}$ and a generation rule
$\mathsf{gen}\stackrel{\beta}{\to} \true +\mathsf{gen}$, 
where in each case the value $\alpha,\beta\in \Lawv$ indicates a \emph{cost} associated with the reduction (e.g.~the estimated clock value for the simulation of each reduction on a given machine model). 
Then, any reduction $\omega: M \to^{*} N$ of a term to (one of its) normal form is  associated with a tropical monomial $P_{\omega}( \alpha,\beta)$ consisting of the sum of the costs of all reductions in $\omega$. For a given normal form $N$, the reductions $\omega_{i}: M \to^{*}N$ give rise to a tLs $\inf_{i\in I}P_{\omega_{i}}(\vec \alpha,\beta)$. 
For example, consider the non-deterministic term below
$$
M :=\mathsf{gen}+  ((\true + \true) + \mathsf{gen})
$$
The (infinitely many) reduction paths leading to $\true$ can be grouped as follows:
\begin{itemize}
\item left, then reduce gen $n+1$-times, then left;
\item right, then left and then either left or right;
\item right twice, then reduce gen $n+1$-times and then left.
\end{itemize}
This leads to the tLs 
$$\varphi_{M\to^{*}\true}(\alpha)=\inf_{n\in \mathbb N}\big\{2\alpha+(n+1)\beta, 3\alpha, 3\alpha+(n+1)\beta\big \}= \min\{ 2\alpha+\beta, 3\alpha\}$$which describes all possible behaviors of $M$. Notice that, since $\alpha$ and $\beta$ are always positive, the power series $\varphi_{M\to^{*}\true}(\alpha)$ is indeed equivalent to the tropical polynomial $\min\{ 2\alpha+\beta, 3\alpha\}$. In other words, of the infinitely many behaviors of $M$, only finitely many have chances to be \emph{optimal}: either left + gen + left, or right + left + (left,right). This seems to be a very interesting feature of tropical semantics: reducing to best-case analysis leads, in many cases, to collapse the infinitary description of \emph{all} behaviors to a purely combinatorial description of the finitely many optimal ones.

Once reduced $\varphi_{M\to^{*}\true}$ to a polynomial, the best behavior among these will depend on the values of $\alpha$ and $\beta$, and by studying the tropical polynomial $\varphi_{M\to^{*}\true}$ one can thus answer questions analogous to 2) above, that is, what are the best choices of costs $\alpha,\beta$ making a \emph{chosen} reduction of $M$ to $\true$ the cheapest one?

>>>>

Let us start with a very basic probabilistic language:
the terms are $M::= \true \mid \false \mid M\oplus_p M$, for $p\in[0,1]$, and the operational semantics is $M\oplus_p N\to pM$ and $M\oplus_p N \to (1-p)N$, so that $M\oplus_p N$ plays the role of a probabilistic coin toss of bias $p$.
Consider the program
$
 M:=(\true \oplus_p\false)\oplus_p((\true\oplus_p\false)\oplus_p(\false\oplus_p\true)).
 $
 Calling $q=1-p$, to each occurrence of $\true$ or $\false$ in $M$, univocally determined by an address
$\omega\in \{l,r\}^{*}$, is associated a monomial $P_{\omega}(p,q)$ which determines the probability of the event ``$M\twoheadrightarrow_{\omega} \true/\false$'', that is, that $M$ reduces to $\true/\false$ according to the choices in $\omega$.
Thinking of $p,q$ as parameters, $P_{\omega}(p,q)$ can thus be read as the \emph{likelihood function} of the event ``$M\twoheadrightarrow \true_\omega/\false_\omega$''.
 For instance, we have
$P_{rll}(p,q):=qp^2$,
$P_{rrr}(p,q):=q^3$, and 
$P_{rrl}(p,q)=P_{rlr}(p,q):=q^2p$.
The polynomial function $Q_{\true}(p,q):=P_{ll}(p,q)+P_{rll}(p,q)+P_{rrr}(p,q)=p^2+p^2q+q^3$ gives instead the probability of the event ``$M\twoheadrightarrow \true$'', and analogously for $Q_{\false}(p,q):=P_{lr}(p,q)+P_{rrl}(p,q)+P_{rlr}(p,q)=pq+2pq^2$.
This way, the probabilistic evaluation of $M$ is presented as a \emph{hidden Markov model} \cite{Baumr966}, a fundamental statistical model, and notably one to which tropical methods are generally applied \cite{Pachter2ll4}.
Typical questions in this case would be, for a fixed $\omega_0$:
%
%The tropical point of view allows now to express two natural questions about this situation:
\begin{enumerate}
 \item Which is the \emph{maximum likelihood estimator} for the event ``$M\twoheadrightarrow \false_{\omega_0}$''?
 I.e., which is the choice of $p,q$ that maximizes the probability $P_{\omega_0}$?
 \item 
Knowing that $M$ produced $\true$ (similarly for $\false$) , which is the \emph{maximum likelihood estimator} for the event ``$M\twoheadrightarrow \false_{\omega_0}$'', knowing that ``$M\twoheadrightarrow \false$''?
I.e., which is the choice of $p,q$ that makes $\omega_0$ the most likely path among those leading to $\false$ (i.e.\ that maximizes the conditional probability $\BB P(``M\twoheadrightarrow \false_{\omega_0}'' \mid ``M\twoheadrightarrow \false'')$)?
\end{enumerate}

@@

This example is inspired from \cite{Manzo2013}. Again, we start from a very basic non-deterministic language with terms $M::= \true \mid  \mathsf{gen} \mid  M + M$, with an operation semantics comprising a non-deterministic reduction rule 
$M_{1}+M_{2} \stackrel{\alpha}{\to} M_{i}$ and a generation rule
$\mathsf{gen}\stackrel{\beta}{\to} \true +\mathsf{gen}$, 
where in each case the value $\alpha,\beta\in \Lawv$ indicates a \emph{cost} associated with the reduction (e.g.~the estimated clock value for the simulation of each reduction on a given machine model). 
Then, any reduction $\omega: M \to^{*} N$ of a term to (one of its) normal form is  associated with a tropical monomial $P_{\omega}( \alpha,\beta)$ consisting of the sum of the costs of all reductions in $\omega$. For a given normal form $N$, the reductions $\omega_{i}: M \twoheadrightarrow N$ give rise to a tLs $\inf_{i\in I}P_{\omega_{i}}(\vec \alpha,\beta)$. 
For example, consider the non-deterministic term
$
M :=\mathsf{gen}+  ((\true + \true) + \mathsf{gen})
$
The (infinitely many) reduction paths leading to $\true$ can be grouped as follows:
\begin{itemize}
\item left, then reduce gen $n+1$-times, then left;
\item right, then left and then either left or right;
\item right twice, then reduce gen $n+1$-times and then left.
\end{itemize}
This leads to the tLs 
$\varphi_{M\twoheadrightarrow \true}(\alpha)=\inf_{n\in \mathbb N}\big\{2\alpha+(n+1)\beta, 3\alpha, 3\alpha+(n+1)\beta\big \}= \min\{ 2\alpha+\beta, 3\alpha\}$, which describes all possible behaviors of $M$. Notice that, since $\alpha$ and $\beta$ are always positive, the power series $\varphi_{M\twoheadrightarrow\true}(\alpha)$ is indeed equivalent to the tropical polynomial $\min\{ 2\alpha+\beta, 3\alpha\}$. In other words, of the infinitely many behaviors of $M$, only finitely many have chances to be \emph{optimal}: either left + gen + left, or right + left + (left,right). This seems to be a very interesting feature of tropical semantics: reducing to best-case analysis leads, in many cases, to collapse the infinitary description of \emph{all} behaviors to a purely combinatorial description of the finitely many optimal ones.

Once reduced $\varphi_{M\twoheadrightarrow\true}$ to a polynomial, the best behavior among these will depend on the values of $\alpha$ and $\beta$, and by studying the tropical polynomial $\varphi_{M\twoheadrightarrow\true}$ one can thus answer questions analogous to 2) above, that is, what are the best choices of costs $\alpha,\beta$ making a \emph{chosen} reduction of $M$ to $\true$ the cheapest one?

### 2023-07-30 CSL24/section3fancy.tex
Commit
Dettagli vari

One can easily check, by induction on a typing derivation, that for any program of $\STLC$ or $\mathrm{PCF}$, the associated matrix is \emph{discrete}, that is, its values are included in $\{0,\infty\}$. Indeed, as suggested in Section \ref{section2}, the actual interest of tropical semantics lies in the interpretation of effectful programs. 
%
%In particular, the \emph{evaluation} morphisms $\RM{ev}$ in $\LREL$ are discrete matrices in $\Lawv^{!(!X\times Y)+X)\times Y}$ given by 
%$\RM{ev}_{[ (\mu,b) ,\mu ] , b }=0$, and being $\infty$ in all other cases, and the coKleisli composition of $s\in\Lawv^{!Y\times Z}$ and $t\in\Lawv^{!X\times Y}$ is the matrix $s\circ_! t\in\Lawv^{!X\times Z}$, $(s\circ_! t)_{\mu,c}:=
%%\begin{align}
%\inf_{n\in\N, b_1\dots,b_n\in Y, \mu = \mu_1+\cdots +\mu_n}
% \left\{s_{[b_1,\dots,b_n],c} + \sum_{i=1}^n t_{\mu_i,b_i}\right\}$.
% For example, setting $\model{A}=\{\star\}$, the interpretation of the $\lambda$-term $\lambda x.\lambda y.xy: (A\to A)\to (A\to A)$ yields the tLs
% $\model{M}^{!}: \Lawv^{\mathbb N}\to \Lawv \to \Lawv$ (recalling that we can identify $\mathcal M_{\mathrm{fin}}(\{\star\})$ with $\mathbb N$) given by 
% $$
% \model{M}^{!}(x)(y)= \RM{ev}^{!}(\langle x,y\rangle)= \inf_{n\in \mathbb N}\big\{ x_{n} + ny\big\}.  
% $$
% Moreover,  
As the homsets $\LREL_{!}(X,Y)$ are $\Lawv$-modules, it is possible to interpret in extensions of $\STLC$ and $\mathrm{PCF}$ comprising $\Lawv$-module operations $\alpha \cdot M$ and $M+N$ \cite{Manzo2013}, by 
letting $\model{\Gamma \vdash \alpha\cdot M:A}=\model{\Gamma \vdash M:A}+\alpha$ and 
$\model{\Gamma \vdash M+N:A}=\min\{\model{\Gamma \vdash M:A},\model{\Gamma \vdash N:A}\}$. 
More precisely, \cite{Manzo2013} considers a language $\mathrm{PCF}^{Q}$ corresponding to $\mathrm{PCF}$ extended with $Q$-module operations, with an operational semantics given by rules $M\stackrel{1}{\to}M'$ for each rule $M\to M'$ of $\mathrm{PCF}$ (here $1$ is the monoidal unit of $Q$) as well as 
$M_{1}+M_{2} \stackrel{1}{\to}M_{i}$ and $\alpha\cdot M \stackrel{\alpha}{\to} M $.
Hence, any reduction $\omega: M\to^{*} N$ is naturally associated with a weight $\mathsf w(\omega)\in Q$.
In particular, from [\cite{Manzo2013}, Theorem V.6] we deduce the following result, which relates the tropical semantics of a program with its best-case operational behavior.  
\begin{proposition}
If $\vdash_{\mathrm{PCF}^{\Lawv}} M:\mathrm{Nat}$, then $\model{\vdash_{\mathrm{PCF}^{\Lawv}} M:\mathrm{Nat}}_{n}=  \inf\big\{ \mathsf w(\omega) \ \big \vert \ \omega : M \to \underline n\big \}$ for all $n\in \mathbb N$.
\end{proposition}

Observe that the two examples shown in Section \ref{section2} can easily be rephrased in the language $\mathrm{PCF}^{\Lawv}$. 
For instance, for the probabilistic example, %for a given valuation $\RM{val}$, one can use a translation %$(M\oplus_{p}N)^{\RM{val}}= \min\{(M^{\RM{val}}+\RM{val}(p)), N^{\RM{val}}+\RM{val}(1-p)\}$, so that the reductions $:M\oplus_{p}N\stackrel{p}{\to} M$ and $M\oplus_{p}N\stackrel{p-1}{\to} N$ translate into a sequence of two reductions $(M\oplus_{p}N)^{\RM{val}} \stackrel{0}{\to} M^{\RM{val}}\stackrel{\RM{val}(p)}{\to} M$ and $(M\oplus_{p}N)^{\RM{val}} \stackrel{0}{\to} N^{\RM{val}}\stackrel{\RM{val}(1-p)}{\to} M$, with the weight of the latter corresponding to the valuation of the weight of the former. 
one can use the translation $(M\oplus_{p}N)^{\circ}= \min\{M^{\circ}+p, N^{\circ}+(1-p)\}$, so that the reductions $:M\oplus_{p}N\stackrel{p}{\to} M$ and $M\oplus_{p}N\stackrel{1-p}{\to} N$ translate into a sequence of two reductions $(M\oplus_{p}N)^{\circ} \stackrel{0}{\to} M^{\circ}\stackrel{p}{\to} M$ and $(M\oplus_{p}N)^{\circ} \stackrel{0}{\to} N^{\circ}\stackrel{1-p}{\to} N^{\circ}$.%, with the weight of the latter corresponding to the valuation of the weight of the former.

>>>>

One can easily check, by induction on a typing derivation, that for any program of $\STLC$ or $\mathrm{PCF}$, the associated matrix is \emph{discrete}, that is, its values are included in $\{0,\infty\}$. Indeed, as suggested in Section \ref{section2}, the actual interest of tropical semantics lies in the interpretation of effectful programs. 
%
%In particular, the \emph{evaluation} morphisms $\RM{ev}$ in $\LREL$ are discrete matrices in $\Lawv^{!(!X\times Y)+X)\times Y}$ given by 
%$\RM{ev}_{[ (\mu,b) ,\mu ] , b }=0$, and being $\infty$ in all other cases, and the coKleisli composition of $s\in\Lawv^{!Y\times Z}$ and $t\in\Lawv^{!X\times Y}$ is the matrix $s\circ_! t\in\Lawv^{!X\times Z}$, $(s\circ_! t)_{\mu,c}:=
%%\begin{align}
%\inf_{n\in\N, b_1\dots,b_n\in Y, \mu = \mu_1+\cdots +\mu_n}
% \left\{s_{[b_1,\dots,b_n],c} + \sum_{i=1}^n t_{\mu_i,b_i}\right\}$.
% For example, setting $\model{A}=\{\star\}$, the interpretation of the $\lambda$-term $\lambda x.\lambda y.xy: (A\to A)\to (A\to A)$ yields the tLs
% $\model{M}^{!}: \Lawv^{\mathbb N}\to \Lawv \to \Lawv$ (recalling that we can identify $\mathcal M_{\mathrm{fin}}(\{\star\})$ with $\mathbb N$) given by 
% $$
% \model{M}^{!}(x)(y)= \RM{ev}^{!}(\langle x,y\rangle)= \inf_{n\in \mathbb N}\big\{ x_{n} + ny\big\}.  
% $$
% Moreover,  
As the homsets $\LREL_{!}(X,Y)$ are $\Lawv$-modules, it is possible to interpret in extensions of $\STLC$ and $\mathrm{PCF}$ comprising $\Lawv$-module operations $\alpha \cdot M$ and $M+N$ \cite{Manzo2013}, by 
letting $\model{\Gamma \vdash \alpha\cdot M:A}=\model{\Gamma \vdash M:A}+\alpha$ and 
$\model{\Gamma \vdash M+N:A}=\min\{\model{\Gamma \vdash M:A},\model{\Gamma \vdash N:A}\}$. 
More precisely, \cite{Manzo2013} considers a language $\mathrm{PCF}^{Q}$ corresponding to $\mathrm{PCF}$ extended with $Q$-module operations, with an operational semantics given by rules $M\stackrel{1}{\to}M'$ for each rule $M\to M'$ of $\mathrm{PCF}$ (here $1$ is the monoidal unit of $Q$) as well as 
$M_{1}+M_{2} \stackrel{1}{\to}M_{i}$ and $\alpha\cdot M \stackrel{\alpha}{\to} M $.
Hence, any reduction $\omega: M\to^{*} N$ is naturally associated with a weight $\mathsf w(\omega)\in Q$.
In particular, from [\cite{Manzo2013}, Theorem V.6] we deduce the following adequation result:
\begin{proposition}
If $\vdash_{\mathrm{PCF}^{\Lawv}} M:\mathrm{Nat}$, then $\model{\vdash_{\mathrm{PCF}^{\Lawv}} M:\mathrm{Nat}}_{n}=  \inf\big\{ \mathsf w(\omega) \ \big \vert \ \omega : M \to \underline n\big \}$ for all $n\in \mathbb N$.
\end{proposition}

The previous result allows to relate the tropical semantics of a program with its best-case operational behavior.  
Observe that the two examples shown in Section \ref{section2} can easily be rephrased in the language $\mathrm{PCF}^{\Lawv}$. 
For instance, for the probabilistic example, %for a given valuation $\RM{val}$, one can use a translation %$(M\oplus_{p}N)^{\RM{val}}= \min\{(M^{\RM{val}}+\RM{val}(p)), N^{\RM{val}}+\RM{val}(1-p)\}$, so that the reductions $:M\oplus_{p}N\stackrel{p}{\to} M$ and $M\oplus_{p}N\stackrel{p-1}{\to} N$ translate into a sequence of two reductions $(M\oplus_{p}N)^{\RM{val}} \stackrel{0}{\to} M^{\RM{val}}\stackrel{\RM{val}(p)}{\to} M$ and $(M\oplus_{p}N)^{\RM{val}} \stackrel{0}{\to} N^{\RM{val}}\stackrel{\RM{val}(1-p)}{\to} M$, with the weight of the latter corresponding to the valuation of the weight of the former. 
one can use the translation $(M\oplus_{p}N)^{\circ}= \min\{M^{\circ}+p, N^{\circ}+(1-p)\}$, so that the reductions $:M\oplus_{p}N\stackrel{p}{\to} M$ and $M\oplus_{p}N\stackrel{1-p}{\to} N$ translate into a sequence of two reductions $(M\oplus_{p}N)^{\circ} \stackrel{0}{\to} M^{\circ}\stackrel{p}{\to} M$ and $(M\oplus_{p}N)^{\circ} \stackrel{0}{\to} N^{\circ}\stackrel{1-p}{\to} N^{\circ}$.%, with the weight of the latter corresponding to the valuation of the weight of the former.

### 2023-07-30 CSL24/section4.tex
Commit
Dettagli vari

In all this section we consider the sets $\Lawv^X$ endowed with the usual sup-norm $\norm{x}_\infty:=\sup_{a\in X} x_a$.

@@

@@

%The norm $\norm{\cdot}_\infty$ naturally induces a metric $\norm{x-y}_{\infty}$ over the spaces $\Lawv^{X}$.
We will show that tLs satisfy suitable Lipschitz properties w.r.t.\ $\norm{\cdot}_\infty$.

Let us first look at tropical linear functions:

\begin{proposition}\label{prop:troplinear}
All tropical \emph{linear} functions $f: \Lawv^{X}\to \Lawv^{Y}$ are non-expansive.  
\end{proposition}
%\begin{proof}[Proof sketch]
%Using the fact that $f(\B x)_{b}= \inf_{a\in X}\matr f_{a,b}+\B x_{a}$,
%the problem reduces to checking that $|(\matr f_{a,b}-\B x_{a})- (\matr f_{a,b}-\B y_{a})| = |\B x_{a}-\B y_{a}|\leq \| \B x-\B y\|_{\infty}$.\end{proof}
This result shows that, in analogy with that happens in usual metric semantics, linear programs are interpreted by non-expansive functions. 
%\begin{proof}
%Using $f(\B x)_{b}= \inf_{a\in X}\matr f_{a,b}+\B x_{a}$,
%first observe that $|(\matr f_{a,b}-\B x_{a})- (\matr f_{a,b}-\B y_{a})| = |\B x_{a}-\B y_{a}|\leq \| \B x-\B y\|_{\infty}$; we now have
%$|f(\B x)_{b}-f(\B y)_{b}| \leq |(\inf_{a\in X}\matr f_{a,b}-\B x_{a})-(\inf_{a\in X}\matr f_{a,b}-\B y_{a})| \leq
%\sup_{a\in X}|(\matr f_{a,b}-\B x_{a})- (\matr f_{a,b}-\B y_{a})|\leq 
% \| \B x-\B y\|_{\infty}$.
%\end{proof}

%Before looking at what happens in the case of non-linear programs, let us make the metric structure of $\LREL$ explicit. 
The following proposition provides a useful characterization of the functional metrics in $\LREL$, relying on 
the bijection between $\HOM{\LREL}{X}{Y}$ and set of tropical linear functions from $\Lawv^{X}$ to $\Lawv^{Y}$.

@@

Let us now look at what happens with tLs, i.e.~when considering the full exponential $!$.
As consequence of Theorem~\ref{theorem:fepsilon}, the tLs with \emph{finitely many} variables are always \emph{locally} Lipschitz on all $\BB R_{>0}$.
%Actually, we can prove a more general statement, also covering the infinitary case.

One could now obtain the locally Lipschitz property of also infinitely many variables tLs from the results in \cite{Cobzas2017}.
In \cite[Theorem 6.4]{Cobzas2017} the locally Lipschitz property is obtained in general for LCTVS under the hypothesis of continuity.
Instead, we replace the continuity hypothesis by the locally bounded condition in our next theorem, which corresponds to the one considered in \cite[Property 6.5]{Cobzas2017}.
Our next theorem is basically a particular case of \cite[Theorem 6.9, Lemma 6.10]{Cobzas2017}, but we prefer instead to state it for a less general case and not talk about more complex argument involving LCTVS.

\begin{theorem}\label{thm:locLip}
 Let $f:V\subseteq (\BB R^X,d) \to (\BB R,\absv \cdot)$, with $V$ open and convex and $d$ any metric.
 If $f$ is concave and locally bounded, then $f$ is locally Lipschitz.
 Moreover, the Lipschitz constant of $f_{\big|_{\overline{B_{\delta}(x)}}}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} \absv f$.
\end{theorem}
\begin{proof}[Proof]
 Call $\overline{B_{\delta}(x)}:=B_1$, $\overline{B_{3\delta}(x)}:=B_3$.
 It suffices to show that for all $x\in V$, there is $\delta>0$ s.t.\ $B_3\subseteq \mathrm{interior}(V)$, $K:=\max_{B_3}  \absv f$ exists and $f_{\big|_{B_1}}$ is $(\frac{1}{\delta}\max_{B_3} \absv f)$-Lipschitz.
A $\delta$ satisfying the first two conditions exists since $V$ is open and bcause $f$ is locally bounded and $B_3$ is compact.
Now fix $y,z\in B_1$.
Call $r:=\frac{d(y,z)}{2\delta}\in[0,1]$.
We have to show that $\absv{f(y)-f(z)}\leq \frac{K}{\delta}d(y,z)=2Kr$.
Wlog $y\neq z$.
So $r\neq 0$ and we can consider $u:=\frac{1+r}{r}z-\frac{1}{r}y$, $v:=\frac{1}{r}y-\frac{r-1}{r}z$.
We have $u,v\in \overline{B_{2\delta}(z)}=:B_2$ (geometrically, those are the intersections between $B_2$ and the segment between $y$ and $z$).
Now we have the convex combinations $z=\frac{1}{1+r}y+\frac{r}{1+r}u$ and $y=(1-r)z+rv$, so the concavity of $f$ entails on one hand:
$f(z)\geq \frac{1}{1+r}f(y)+\frac{r}{1+r}f(z)\geq \frac{f(y)}{1+r} - \frac{rK}{1+r}$, i.e.\ $f(y)-f(z)\leq r(K+f(z))\leq 2rK$, and on the other hand:
$f(y)\geq (1-r)f(z)+rf(v)\geq f(z)-r(f(z)+K)$, i.e.\ $f(z)-f(y)\leq r(f(z)+K)\leq 2rK$.
The last inequalities follow because $B_2\subseteq B_3$.
Putting those together, we have the thesis.
\end{proof}

Now since $(0,\infty)^X$ is open and convex in $\BB R^X$, %and all tLs are non-negative, 
we immediately have:

\begin{theorem}\label{thmTLSlocLip}
 All tLs $f:\Lawv^X\to\Lawv$ are locally Lipschitz on $(0,\infty)^X$.
 Moreover, the Lipschitz constant of $f_{\big|_{\overline{B_{\delta}(x)}}}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} f$.
\end{theorem}

>>>>

@@

In all the remainder of this section we consider the sets $\Lawv^X$ endowed with the usual sup-norm $\norm{x}_\infty:=\sup_{a\in X} x_a$.

@@

%The norm $\norm{\cdot}_\infty$ naturally induces a metric $\norm{x-y}_{\infty}$ over the spaces $\Lawv^{X}$.
%We will show that tLs satisfy suitable Lipschitz properties w.r.t.\ $\norm{\cdot}_\infty$.

Let us first look at tropical linear functions:
in analogy with that happens in usual metric semantics, we have:

\begin{proposition}\label{prop:troplinear}
All tropical \emph{linear} functions $f: \Lawv^{X}\to \Lawv^{Y}$ are non-expansive (i.e.\ $1$-Lipschitz).  
\end{proposition}
%\begin{proof}[Proof sketch]
%Using the fact that $f(\B x)_{b}= \inf_{a\in X}\matr f_{a,b}+\B x_{a}$,
%the problem reduces to checking that $|(\matr f_{a,b}-\B x_{a})- (\matr f_{a,b}-\B y_{a})| = |\B x_{a}-\B y_{a}|\leq \| \B x-\B y\|_{\infty}$.\end{proof} 
%\begin{proof}
%Using $f(\B x)_{b}= \inf_{a\in X}\matr f_{a,b}+\B x_{a}$,
%first observe that $|(\matr f_{a,b}-\B x_{a})- (\matr f_{a,b}-\B y_{a})| = |\B x_{a}-\B y_{a}|\leq \| \B x-\B y\|_{\infty}$; we now have
%$|f(\B x)_{b}-f(\B y)_{b}| \leq |(\inf_{a\in X}\matr f_{a,b}-\B x_{a})-(\inf_{a\in X}\matr f_{a,b}-\B y_{a})| \leq
%\sup_{a\in X}|(\matr f_{a,b}-\B x_{a})- (\matr f_{a,b}-\B y_{a})|\leq 
% \| \B x-\B y\|_{\infty}$.
%\end{proof}

%Before looking at what happens in the case of non-linear programs, let us make the metric structure of $\LREL$ explicit. 
The following proposition provides a useful characterization of the functional metrics in $\LREL$, relying on 
the bijection between $\HOM{\LREL}{X}{Y}$ and set of tropical linear functions from $\Lawv^{X}$ to $\Lawv^{Y}$.
Remember that $d_\infty(f,g):=\sup_{a\in X} d_\infty(f(x),f(y))$.

@@

Let us now look at what happens with tLs, i.e.~when considering the full exponential $!$.
As consequence of Theorem~\ref{theorem:fepsilon}, the tLs with \emph{finitely many} variables are always \emph{locally} Lipschitz on all $(0,\infty)^X$.
%Actually, we can prove a more general statement, also covering the infinitary case.

We wish now to move to the case of infinitely many variables.
In \cite[Theorem 6.4]{Cobzas2017} a locally Lipschitz property is obtained for locally convex topological vector spaces, under the hypothesis of continuity. \cite[Proposition 6.5]{Cobzas2017} shows that continuity is used in order to have a locally bounded condition, the true main ingredient of the proof.
Instead of showing how our case fits into such theorems, we prefer to state a theorem, which is basically a particular case of \cite[Theorem 6.9, Lemma 6.10]{Cobzas2017}, already fitting in the topics under consideration in this paper.

\begin{theorem}\label{thmTLSlocLip}
 Let $f:V\subseteq (\BB R^X,d) \to (\BB R,\absv \cdot)$, with $V$ open and convex and $d$ any metric.
 If $f$ is concave and locally bounded, then $f$ is locally Lipschitz.
 Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} \absv f$.

Therefore, since $(0,\infty)^X$ is open and convex in $(\BB R^X,\norm\cdot)$, %and all tLs are non-negative, 
we immediately have that all tLs $f:\Lawv^X\to\Lawv$ are locally Lipschitz on $(0,\infty)^X$.
Moreover, the Lipschitz constant of $f$ on $\overline{B_{\delta}(x)}$ can be chosen to be $\frac{1}{\delta}\max_{\overline{B_{3\delta}(x)}} f$.
\end{theorem}
%\begin{proof}[Proof]
% Call $\overline{B_{\delta}(x)}:=B_1$, $\overline{B_{3\delta}(x)}:=B_3$.
% It suffices to show that for all $x\in V$, there is $\delta>0$ s.t.\ $B_3\subseteq \mathrm{interior}(V)$, $K:=\max_{B_3}  \absv f$ exists and $f_{\big|_{B_1}}$ is $(\frac{1}{\delta}\max_{B_3} \absv f)$-Lipschitz.
%A $\delta$ satisfying the first two conditions exists since $V$ is open and bcause $f$ is locally bounded and $B_3$ is compact.
%Now fix $y,z\in B_1$.
%Call $r:=\frac{d(y,z)}{2\delta}\in[0,1]$.
%We have to show that $\absv{f(y)-f(z)}\leq \frac{K}{\delta}d(y,z)=2Kr$.
%Wlog $y\neq z$.
%So $r\neq 0$ and we can consider $u:=\frac{1+r}{r}z-\frac{1}{r}y$, $v:=\frac{1}{r}y-\frac{r-1}{r}z$.
%We have $u,v\in \overline{B_{2\delta}(z)}=:B_2$ (geometrically, those are the intersections between $B_2$ and the segment between $y$ and $z$).
%Now we have the convex combinations $z=\frac{1}{1+r}y+\frac{r}{1+r}u$ and $y=(1-r)z+rv$, so the concavity of $f$ entails on one hand:
%$f(z)\geq \frac{1}{1+r}f(y)+\frac{r}{1+r}f(z)\geq \frac{f(y)}{1+r} - \frac{rK}{1+r}$, i.e.\ $f(y)-f(z)\leq r(K+f(z))\leq 2rK$, and on the other hand:
%$f(y)\geq (1-r)f(z)+rf(v)\geq f(z)-r(f(z)+K)$, i.e.\ $f(z)-f(y)\leq r(f(z)+K)\leq 2rK$.
%The last inequalities follow because $B_2\subseteq B_3$.
%Putting those together, we have the thesis.
%\end{proof}
